\bf\mftitle

### 2011-01-12 doc/jInfer/full/architecture/architecture.tex
3154619 Create documentation for jInfer architecture - better fonts and pictures
3154619 Create documentation for jInfer architecture - minors
3154619 Create documentation for jInfer architecture - verbatim with spaces instead tabs
Documentation: Architecture: Inference Process

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}

@@

Regular expression is in fact $n$-ary tree, for example expression $(a, b, ((c | d), e), f)$ can be viewed as in fig. \ref{reg_tree}.
\begin{figure}
\caption{Example tree for regular expression $(a, b, ((c | d), e), f)$} \label{reg_tree}
\centering\includegraphics[scale=0.5]{reg_tree}
\end{figure}
We implement this tree by member of \code{Regexp} class called \code{children}, which is of type \code{List<Regexp<T>{}>}.
List contains children of regexp in means of regexp tree.

@@

The first stands for bare node in XML document tree, it has its name and context withing the tree (path from root).
The latter two extends \code{NamedNode} interface.
\code{StructuralNode} is for nodes, which form structure of XML document tree: elements and text nodes.
\code{ContentNode} is for nodes, that have content in XML documents: text nodes and attributes.
We have three classes: \code{Element} for elements, \code{SimpleData} for text nodes, \code{Attribute} for attributes (see package \code{cz.cuni.mff.ksi.jinfer.base.objects.nodes}).
In theory, the classes and interfaces would be layed out as on fig. \ref{interfaces_nodes}
\begin{figure}
\caption{How should interfaces and classes for XML representation look like in theory} \label{interfaces_nodes}
\centering\includegraphics[scale=0.5]{interfaces_nodes}
\end{figure}
\begin{figure}
\caption{How are interfaces and classes for XML representation arranged in practice} \label{nodes}
\centering\includegraphics[scale=0.5]{nodes}
\end{figure}
\begin{figure}
\caption{XML representing interfaces and classes in detail} \label{nodes_full}
\centering\includegraphics[scale=0.3]{nodes_full}
\end{figure}

@@

% TODO explain what IG (or a grammar, for that matter) is
% TODO explain metadata and those that we already use
% TODO section about NBMs

\section{Inference process}
\begin{figure}
	\caption{High-level view of the inference process} \label{inference_process}
	\centering\includegraphics[scale=0.5]{inference_process}
\end{figure}

@@

The process by which jInfer infers the resulting schema from various inputs (inference process) is summarized by fig. \ref{inference_process}. From the high-level viewpoint, it consists of three consecutive steps carried out by three different modules:
\begin{enumerate}
	\item Initial Grammar (IG) generation: done by the \textit{Initial Grammar Generator (IGG)} module, this is the process of converting all of the inputs to IG representation. All documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step.\\
	% TODO anti: how about the whole examples in italic?
	For example, a trivial XML document
	% TODO anti: please format the indentation	
	\begin{verbatim}
		<a>
			<b>
				<c/>
			</b>
			<d/>
		</a>
	\end{verbatim}
	would translate into the following IG rules
	\begin{verbatim}
		a -> b, d
		b -> c
	\end{verbatim}	
	\item Simplification: done by the \textit{Simplifier} module, this is the process of simplifying, compressing or somehow compactly describing the IG by a smaller number of (more complex) rules. User interaction might be used in this step to help achieve better simplification. At the end of this step, all rules are sent to the last step.\\
	For example, rules
	\begin{verbatim}
		a -> b
		a -> c, d, d, d
	\end{verbatim}
	could be simplified to a single rule
	\begin{verbatim}
		a -> b | (c, d*)
	\end{verbatim}
	\item Schema export: done by the \textit{Schema Generator (SchemaGen)} module, this is the process of actually creating the resulting schema from the simplified rules. Result of this step is a string representation of the schema, which is sent back to the framework (and later displayed, saved, etc).\\
	% TODO vektor Example
\end{enumerate}
Important thing to note here is that all these steps are executed consecutively. That means, \textit{Simplifier} is only started \textit{after} the \textit{IGG} completely finished its work and returned IG to be simplified. Similarly, \textit{SchemaGen} gets all the rules to export at once, in one list.\\

>>>>

\documentclass[a4paper,10pt,oneside]{article}
\usepackage{graphicx}
\usepackage{color}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tgpagella}
%\usepackage[scale=0.9]{tgcursor}
%\usepackage[scale=0.9]{tgheros}

@@

Regular expression is in fact $n$-ary tree, for example expression $(a, b, ((c | d), e), f)$ can be viewed as in fig. \ref{reg_tree}.
\begin{figure}
\caption{Example tree for regular expression $(a, b, ((c | d), e), f)$} \label{reg_tree}
\centering\includegraphics[scale=\myscale]{reg_tree}
\end{figure}
We implement this tree by member of \code{Regexp} class called \code{children}, which is of type \code{List<Regexp<T>{}>}.
List contains children of regexp in means of regexp tree.

@@

The first stands for bare node in XML document tree, it has its name and context withing the tree (path from root).
The latter two extends \code{NamedNode} interface.
\code{StructuralNode} is for nodes, which form structure of XML document tree: elements and text nodes.
\code{ContentNode} is for nodes, that have content in XML documents: text nodes and attributes.
We have three classes: \code{Element} for elements, \code{SimpleData} for text nodes, \code{Attribute} for attributes (see package \code{cz.cuni.mff.ksi.jinfer.base.objects.nodes}).
In theory, the classes and interfaces would be layed out as on fig. \ref{interfaces_nodes}
\begin{figure}
\caption{How should interfaces and classes for XML representation look like in theory} \label{interfaces_nodes}
\centering\includegraphics[scale=\myscale]{interfaces_nodes}
\end{figure}
\begin{figure}
\caption{How are interfaces and classes for XML representation arranged in practice} \label{nodes}
\centering\includegraphics[scale=\myscale]{nodes}
\end{figure}
\begin{figure}
\caption{XML representing interfaces and classes in detail} \label{nodes_full}
\centering\includegraphics[scale=0.3]{nodes_full}
\end{figure}

@@

\subsection{Rules, Grammars, etc}

jInfer and its documentation uses extended context-free grammars. % TODO vektor Cite, for example http://www.engr.mun.ca/~theo/Courses/fm/pub/context-free.pdf
\textit{Rules} in such grammar are in the form
% TODO anti Please make a proper arrow
\[  Left Hand Side (LHS)  -> Right Hand Side (RHS) \]
where LHS is a letter of the alphabet (token), RHS is a regular expression over this alphabet. Example would be
\begin{verbatim}
  a -> b, (c | d)*
\end{verbatim}

In jInfer each such rule is represented with an \code{Element} instance. In this representation, the \code{Element} itself is the LHS, its \code{subnodes} are the RHS.\\

Another important notion is a \textit{grammar}. A grammar consists of its rules, so in jInfer a grammar is just a collection of \code{Element}s. Closely related term is \textit{Initial Grammar}, which for us is a grammar consisting of rules with \textit{simple} right hand sides, i.e. just concatenations of tokens. Initial Grammar is produced by Initial Grammar Generator. % TODO vektor Link

% TODO explain metadata and those that we already use
% TODO section about NBMs

\section{Inference process}
\begin{figure}
	\caption{High-level view of the inference process} \label{inference_process}
	\centering\includegraphics[scale=\myscale]{inference_process}
\end{figure}

@@

The process by which jInfer infers the resulting schema from various inputs (inference process) is summarized by fig. \ref{inference_process}. From the high-level viewpoint, it consists of three consecutive steps carried out by three different modules:
\begin{enumerate}
	\item Initial Grammar (IG) generation: done by the \textit{Initial Grammar Generator (IGG)} module, this is the process of converting all of the inputs to IG representation. All documents, schemas and queries selected as input are evaluated, simple rules are extracted and in the end sent to the next step.\\
	% TODO anti: how about the whole examples in italic?
	For example, a trivial XML document
	\begin{verbatim}
  <a>
    <b>
      <c/>
    </b>
    <d/>
  </a>
	\end{verbatim}
	would translate into the following IG rules
	\begin{verbatim}
		a -> b, d
		b -> c
	\end{verbatim}	
	\item Simplification: done by the \textit{Simplifier} module, this is the process of simplifying, compressing or somehow compactly describing the IG by a smaller number of (more complex) rules. User interaction might be used in this step to help achieve better simplification. At the end of this step, all rules are sent to the last step.\\
	For example, rules
	\begin{verbatim}
		a -> b
		a -> c, d, d, d
	\end{verbatim}
	could be simplified to a single rule
	\begin{verbatim}
		a -> b | (c, d*)
	\end{verbatim}
	\item Schema export: done by the \textit{Schema Generator (SchemaGen)} module, this is the process of actually creating the resulting schema from the simplified rules. Result of this step is a string representation of the schema, which is sent back to the framework (and later displayed, saved, etc).\\
	% TODO vektor Example
\end{enumerate}
Important thing to note here is that all these steps are executed consecutively. That means, \textit{Simplifier} is only started \textit{after} the \textit{IGG} completely finished its work and returned IG to be simplified. Similarly, \textit{SchemaGen} gets all the rules to export at once, in one list.\\
